Why robot programming is hard

### 2013-07-18 paper.tex
Checkpoint.
Folded in comments from Kaijen and Andreas. Starting to make the message come out more strongly in the intro and case study.

\documentclass{article}
\usepackage{graphicx}

@@

I never thought I would be a roboticist. In grad school, our robotics team spent months teaching little robot dogs to play soccer and solving other seemingly simple problems. Far better to work in the world of software, I figured, where the systems I built could make a real difference in people's lives. Software robots would automate repetitive tasks in computer use and free people from mundane work. End user programming would enable millions of ordinary computer users to customize and adapt systems to their own needs.

But that world is changing. Computing has moved off the desktop and into the smartphones, smart environments, and the world aronud us. Innovation is happening in the real world, using computing to effect physical change in our environment. The past decades have seen industrial robots increase manufacturing efficiency on the factory floor. Even more, we are now seeing robots tackle tasks from our daily lives. Robotic vacuum cleaners are changing the way we clean our homes. Robotic lawnmowers keep the yard tidy without having to lift a finger. Nest's robotic thermostat learns your habits and keeps your home at the optimal temperature. Google's driverless car could change the way we commute.

@@

Yet the holy grail of personal robotics remains elusive. When will we have our own Rosie the Robot (from the Jetsons cartoon) to take over housekeeping chores? When will disabled adults have robot helpers to assist them in activities of daily living? All of these advances in robotics still fall short of the goal of creating personal robots that can do whatever we want them to do.

Actually, the problem is not that robots {\em cannot} do what we want them to do. The problem is that it is too difficult to {\em tell} robots what we want them to do. The first company to crack the nut of end user robot programming will corner the market on household robots. Human-robot interaction is the missing link for widespread personal robotics.

@@

One potential use case for personal robotics will be in-home care and assistance for disabled adults. The Robots for Humanity project~\cite{rfh}, a collaboration between Willow Garage and Georgia Tech, aims to develop technology that enables persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes have enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Limited to the ability to move a single finger, Henry has been able to instruct Willow Garage's PR2 robots to open cabinets, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

Yet our experience with Robots for Humanity underscores how difficult it is for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple tasks such as opening a cabinet door could take Henry upwards of 15 minutes to accomplish.  While Henry has only limited interaction capabilities due to motor impairment, he also has incredible motivation. Ordinary consumers, on the other hand, while fully able-bodied, may not share the same perseverance in programming robots.

@@

The Robot Operating System\cite{ros} is one of the software advances that supports the rapid development of new robotics technology.  First developed in 2007 at the Stanford Articifial Intelligence Laboratory, ROS was maintained by Willow Garage from 2008-2013 and has recently transitioned into the stewardship of the non-profit Open Source Robotics Foundation.

@@

This interface, while well suited for roboticists who need to examine the low-level sensor data and messages being processed by the robot, is not very user-friendly. It also illustrates some of the difficulties inherent in robot programming, which I will detail next.

@@

High dimensionality is even more important in specifying how to grasp and place objects in the world. A solid object has six degrees of freedom: three degrees to specify its position, and another three to specify its rotation. The human arm, and the most advanced robotic arms, have seven degrees of freedom.  Many robot arm control interfaces ask users to independently specify joint angles for each of those seven degrees of freedom. These interfaces are cumbersome because the joints are interdependent: setting the angle of one joint, such as the elbow, dictates what angles are possible for the shoulder and vice versa.

Expert roboticists manage high-dimensionality by representing the world in terms of {\em reference frames} (coordinate systems that are defined relative to particular robot part, such as the base or the head-mounted camera) and {\em quaternions} (a matrix that represents position and orientation in 3D space relative to a reference frame). Robot joint positions are typically specified relative to their own reference frame; for example, a robot's wrist, which is located at the end of its arm, can be commanded to rotate around the axis that is aligned with the robot's forearm (similar to the motion we use to screw in a light bulb). Calculating the absolute position of a robot's hand (for example, to determine whether it is near the door handle) requires a series of complex matrix calculations starting from the position of its base, through its shoulder, and all the joints, down to the hand. The opposite problem, calculating the joint positions required to get the hand to a particular position, is called {\em inverse kinematics} and is typically solved using heuristic approximation methods. More generally, techniques for {\em motion planning} calculate whole trajectories of how to move an arm from a starting location to an ending location, using inverse kinematics to compute trajectories for each of the joints in turn.

@@

In addition, robot hardware itself often produces unpredictable results. Unlike software, mechanical parts can slip, burn out, or break. Giving a robot the same instruction multiple times may or may not result in the same outcome each time.

What does this mean for robot programmers? Robot programs must have mechanisms in place to deal with uncertain truth. Rather than simplying instructing the robot to pick up an object, the program must specify what to do if the object is not {\em where} it thinks it is, is not {\em what} it thinks it is, or if the robot fails to perform the desired grasp.

>>>>

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}

@@

I never thought I would be a roboticist. In grad school, our robotics team spent months teaching little robot dogs to play soccer and solving other seemingly simple problems. Far better to work in the world of software, I figured, where the systems I built could make a real difference in people's lives. Software could automate repetitive tasks in computer use and free people from mundane work. My chosen field, end user programming, had the potential to enable millions of ordinary computer users to customize and adapt systems to their own needs.

But the world is changing. Computing has moved off the desktop and into the smartphones, smart environments, and the world around us. Innovation is happening in the physical world, using computing to effect material change in our environment. The past decades have seen industrial robots increase manufacturing efficiency on the factory floor. Even more, we are now seeing robots tackle tasks from our daily lives. Robotic vacuum cleaners are changing the way we clean our homes. Robotic lawnmowers keep the yard tidy without having to lift a finger. Nest's robotic thermostat learns your habits and keeps your home at the optimal temperature. Google's driverless car could change the way we commute.

@@

Yet the holy grail of personal robotics remains elusive. When will we have our own Rosie the Robot to take over housekeeping chores? When will disabled adults have robot helpers to assist them in activities of daily living? All of these advances in robotics still fall short of the goal of creating personal robots that can do whatever we want them to do.

What I have found is that the problem is not that robots {\em cannot} do what we want them to do. The problem is that it is too difficult to {\em tell} robots what we want them to do. Robot control can be roughly categorized into three distinct phases: direct demonstration, tele-operation (or teleop for short), and higher-level programming.

In demonstrational interfaces, a human operator moves a robot through a series of positions to demonstrate the motion it should perform. The robot then continues to perform this motion unattended. This style of programming is common for industrial robots, which can be programmed to move items around on conveyor belts or do precise assembly tasks. Demonstrational interfaces are typically limited to repeating only what was previously demonstrated, in the worst case performing the exact same motion down to the millimeter regardless of whether the part is in the same place as it was last time.

Teleop consists of synchronously controlling a robot using software and hardware interfaces from a computer which may or may not be co-located with the robot. For example, a PS3 joystick controller could be used to drive a robot around a building, and activate certain pre-programmed motions such as turning its head or extending its arms. Remote teleop consists of controlling the robot over a distance, perhaps over the internet, using GUIs that allow the human operator to move various parts of the robot.

Both demonstration and teleop suffer the limitation that a human operator must be immediately in the loop while the robot is operating. To make robots useful in the home, they will need to function {\em autonomously}, which leads to the notion of robot programming: giving high-level directives to robots so that they can function on their own without direct human guidance.  A full programming solution will likely involve a blend of different techniques including demonstration, teleop, and autonomous subroutines. For example, a task of setting the dinner table may include a high-level program of fetching silverware from the kitchen and placing it on the dining table, a demonstration of how to open this particular style of cabinet, and teleoperation when something goes awry (such as the silverware drawer being empty).

For personal robotics to truly succeed, robot programming will need to be simple enough that non-experts can do it.

@@

One important use case for personal robotics will be in-home care and assistance for disabled adults. The Robots for Humanity project~\cite{rfh}, a collaboration between Willow Garage and Georgia Tech, aims to develop technology that enables persons with motor impairments to perform personal care tasks that they otherwise might not have been able to perform without assistance, such as picking up a dropped object, fetching a towel, or even scratching an itch. Its first prototypes have enabled Henry Evans, a mute quadriplegic, to act in the world for the first time since a brainstem stroke left him paralyzed. Through physical therapy, Henry has regained the ability to make limited head motions (enough to control a cursor with his gaze) and to click a button with one finger.  Using a teleop interface custom designed for him, Henry has been able to instruct Willow Garage's PR2 robots to open cabinets in his home, fetch food from the refrigerator, and even dole out candy to trick-or-treating kids on his behalf.

% TODO: conflating robot programming with teleop

Yet our experience with Robots for Humanity underscores how difficult it is for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door could take Henry upwards of 15 minutes to accomplish.  While Henry has only limited interaction capabilities due to motor impairment, he is also strongly motivated. Ordinary able-bodied consumers, on the other hand, would quickly lose patience. If basic robot control via teleoperation is so time-consuming, what hope do we have that people will ever get around to the general task of robot programming?

Granted, designing programming interfaces for motor impaired users is like playing the piano with both hands tied behind your back. Yet while Henry's case highlights the challenges of creating usable interfaces for assistive robotics, it also showcases the enormous benefit to humanity of enabling disabled adults to regain some of their independence.

@@

The Robot Operating System\cite{ros} is one of the software advances that supports the rapid development of new robotics technology.  First developed in 2007 at the Stanford Artificial Intelligence Laboratory, ROS was maintained by Willow Garage from 2008-2013 and has recently transitioned into the stewardship of the non-profit Open Source Robotics Foundation.

@@

This interface, while well suited for roboticists who need to examine the low-level sensor data and messages being processed by the robot, is not very user-friendly. It also illustrates some of the difficulties inherent in robot programming, described below.

@@

High dimensionality is even more important in specifying how to grasp and place objects in the world.  In many cases, the robot itself might occlude the object it is trying to manipulate. For example, if a robot uses its arm to pick up an object on a table in front of it, the arm might get in the way of a sensor mounted on its head, preventing the robot from seeing the exact thing it is trying to grasp! One solution developed by roboticists has been to include additional cameras on different parts of the robot, such as a forearm-mounted camera, to get a better view on the object being picked up. However, making sense of multiple translated and rotated video streams can be confusing for those of us accustomed to seeing only from our own head-mounted eye sensors!

High dimensionality also makes it difficult to specify to how pick up objects using a robotic gripper. A solid object has six degrees of freedom: three degrees to specify its position, and another three to specify its rotation. The human arm, and the most advanced robotic arms, have seven degrees of freedom.  Many robot arm control interfaces ask users to independently specify joint angles for each of those seven degrees of freedom. These interfaces are cumbersome because the joints are interdependent: setting the angle of one joint, such as the elbow, dictates what angles are possible for the shoulder and vice versa. There is a tradeoff between exposing these joint limitations to the programmer, which allows precise yet complex control, versus insulating a programmer from those considerations, which may lead to unpredictable behavior that is difficult to debug.

Expert roboticists manage high-dimensionality by representing the world in terms of {\em reference frames} (coordinate systems that are defined relative to particular robot part, such as the base or the head-mounted camera) and use matrix algebra to convert between different frames of reference. Robot joint positions are typically specified relative to their own reference frame; for example, a robot's wrist, which is located at the end of its arm, can be commanded to rotate around the axis that is aligned with the robot's forearm (similar to the motion we use to screw in a light bulb). Calculating the absolute position of a robot's hand (for example, to determine whether it is near the door handle) requires a series of complex matrix calculations starting from the position of its base, through its shoulder, and all the joints, down to the hand. The opposite problem, calculating the joint positions required to get the hand to a particular position, is called {\em inverse kinematics} and is typically solved using either analytical equations or heuristic approximation methods.

% TODO: introduce DOF acronym earlier

@@

In addition, robot hardware itself often produces unpredictable results. Unlike software, mechanical parts can slip, burn out, or break. They take time to actuate and do not come to a stop immediately. Giving a robot the same instruction multiple times may or may not result in the same outcome each time.

What does this mean for robot programmers? Robot programs must have mechanisms in place to deal with uncertain truth and unpredictable hardware. Robot control loops usually sense the physical hardware during control to determine whether a motion was carried out successfully. 
% TODO
Rather than simplying instructing the robot to pick up an object, the program must specify what to do if the object is not {\em where} it thinks it is, is not {\em what} it thinks it is, or if the robot fails to perform the desired grasp.
